\bf Model Visualization Techniques for a Social Network Model

### 2017-05-31 paper.Rnw
finish intro

\par The guiding principles of model visualization introduced in \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, these steps are largely kept in the background of the estimation process. With the help of static and dynamic visualizations we bring these elements and others into the foreground, leading to a better understanding and higher accessibility of stochastic actor-oriented models  for social network analysts.

In Section~\ref{sec:netintro}, we introduce the concepts of networks and network visualizations. In Section~\ref{sec:saoms}, we present the SAOM for social network analysis. Then, in Section~\ref{sec:ModelVis}, we combine concepts from Sections~\ref{sec:netintro} and \ref{sec:saoms} in an application of the model-vis paradigm, and conclude with a discussion in Section~\ref{sec:discussion}.

>>>>

\par The guiding principles of model visualization introduced in \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out of the black box and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which is attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks on these estimates rely on simulations from the fitted model. Again, these steps are largely kept in the background of the estimation process. With the help of static and dynamic visualizations we bring these elements and others into the foreground, leading to a better understanding and higher accessibility of stochastic actor-oriented models  for social network analysts.

In Section~\ref{sec:netintro}, we introduce the concepts of networks and network visualizations. In Section~\ref{sec:saoms}, we present the SAOM for social network analysis. In Section~\ref{sec:ModelVis}, we combine concepts from Sections~\ref{sec:netintro} and \ref{sec:saoms} in an application of the model-vis paradigm, and conclude with a discussion in Section~\ref{sec:discussion}.
